---
title: "p8105_hw6_yl5219"
author: "Yuqing Liu"
date: "2023-11-30"
output: github_document
---


```{r setup, include=FALSE}
library(tidyverse)
library(viridis)
library(modelr)
library(ggplot2)
library(purrr)

knitr::opts_chunk$set(
	echo = TRUE,
	warning = FALSE,
	fig.width = 8, 
  fig.height = 6,
  out.width = "90%"
)

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d

theme_set(theme_minimal() + theme(legend.position = "bottom"))
```

### Problem 1

In the data cleaning code below we create a `city_state` variable, change `victim_age` to numeric, modifiy victim_race to have categories white and non-white, with white as the reference category, and create a `resolution` variable indicating whether the homicide is solved. Lastly, we filtered out the following cities: Tulsa, AL; Dallas, TX; Phoenix, AZ; and Kansas City, MO; and we retained only the variables `city_state`, `resolution`, `victim_age`, `victim_sex`, and `victim_race`.

```{r q1_data_cleaning}
homicide_df = 
  read_csv("./homicide-data.csv", na = c("", "NA", "Unknown")) |> 
  mutate(
    city_state = str_c(city, state, sep = ", "),
    victim_age = as.numeric(victim_age),
    resolution = case_when(
      disposition == "Closed without arrest" ~ 0,
      disposition == "Open/No arrest"        ~ 0,
      disposition == "Closed by arrest"      ~ 1)
  ) |> 
  filter(victim_race %in% c("White", "Black")) |> 
  filter(!(city_state %in% c("Tulsa, AL", "Dallas, TX", "Phoenix, AZ", "Kansas City, MO"))) |> 
  select(city_state, resolution, victim_age, victim_sex, victim_race)
```

Next we fit a logistic regression model using only data from Baltimore, MD. We model `resolved` as the outcome and `victim_age`, `victim_sex`, and `victim_race` as predictors. We save the output as `baltimore_glm` so that we can apply `broom::tidy` to this object and obtain the estimate and confidence interval of the adjusted odds ratio for solving homicides comparing non-white victims to white victims.

```{r q1_glm_baltimore}
baltimore_glm = 
  filter(homicide_df, city_state == "Baltimore, MD") |> 
  glm(resolution ~ victim_age + victim_sex + victim_race, family = binomial(), data = _)

baltimore_glm |> 
  broom::tidy() |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(OR, OR_CI_lower, OR_CI_upper) |>
  knitr::kable(digits = 3)
```

Below, by incorporating `nest()`, `map()`, and `unnest()` into the preceding Baltimore-specific code, we fit a model for each of the cities, and extract the adjusted odds ratio (and CI) for solving homicides comparing non-white victims to white victims. We show the first 5 rows of the resulting dataframe of model results.

```{r q1_glm_all_cities}
model_results = 
  homicide_df |> 
  nest(data = -city_state) |> 
  mutate(
    models = map(data, \(df) glm(resolution ~ victim_age + victim_sex + victim_race, 
                             family = binomial(), data = df)),
    tidy_models = map(models, broom::tidy)) |> 
  select(-models, -data) |> 
  unnest(cols = tidy_models) |> 
  mutate(
    OR = exp(estimate), 
    OR_CI_upper = exp(estimate + 1.96 * std.error),
    OR_CI_lower = exp(estimate - 1.96 * std.error)) |> 
  filter(term == "victim_sexMale") |> 
  select(city_state, OR, OR_CI_lower, OR_CI_upper)

model_results |>
  slice(1:5) |> 
  knitr::kable(digits = 3)
```

Below we generate a plot of the estimated ORs and CIs for each city, ordered by magnitude of the OR from smallest to largest. From this plot we see that most cities have odds ratios that are smaller than 1, suggesting that crimes with male victims have smaller odds of resolution compared to crimes with female victims after adjusting for victim age and race. This disparity is strongest in New yrok. In roughly half of these cities, confidence intervals are narrow and do not contain 1, suggesting a significant difference in resolution rates by sex after adjustment for victim age and race. 

```{r q1_plot}
model_results |> 
  mutate(city_state = fct_reorder(city_state, OR)) |> 
  ggplot(aes(x = city_state, y = OR)) + 
  geom_point() + 
  geom_errorbar(aes(ymin = OR_CI_lower, ymax = OR_CI_upper)) + 
  theme(axis.text.x = element_text(angle = 90, hjust = 1))
```

### Problem 2

**Import dataset**
```{r}
weather_df = 
  rnoaa::meteo_pull_monitors(
    c("USW00094728"),
    var = c("PRCP", "TMIN", "TMAX"), 
    date_min = "2022-01-01",
    date_max = "2022-12-31") |>
  mutate(
    name = recode(id, USW00094728 = "CentralPark_NY"),
    tmin = tmin / 10,
    tmax = tmax / 10) |>
  select(name, id, everything())
```

**Build a regression model for `tmax`**

```{r}
model_tmax <- lm(tmax ~ tmin + prcp, data = weather_df)
summary(model_tmax)
```

**Perform bootstrap analysis for $\hat r^2$ and $\log(\hat\beta_1\times \hat\beta_2)$**
**Set seed**
```{r}
set.seed(123)
```

**Write a function to generate 5000 bootstraps samples**

```{r}
boot_sample = function(df) {
  sample_frac(df, replace = TRUE)
}
```


**Generate 5000 bootstraps samples**
```{r}
boot_straps = 
  tibble(strap_number = 1:5000) |> 
  mutate(
    strap_sample = map(strap_number, \(i) boot_sample(df = weather_df))
  )

boot_straps
```

**Combine 5000 bootstraps results**

```{r}
bootstrap_results = 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::tidy)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 

```


**Compute the log of the product of beta estimates $\log(\hat\beta_1\times \hat\beta_2)$ of 5000 bootstraps**

```{r}
log_betas <-  
  bootstrap_results %>%
  group_by(strap_number) %>%
  summarise(log_betas = log(estimate[2] * estimate[3])) %>%
  select(log_betas, strap_number)
```


**Generate another 5000 bootstrap estimates**
```{r}
bootstrap_results2 <- 
  boot_straps %>% 
  mutate(
    models = map(strap_sample, ~lm(tmax ~ tmin + prcp, data = .x) ),
    results = map(models, broom::glance)) %>% 
  select(-strap_sample, -models) %>% 
  unnest(results) 
```

**Compute the R-Squared $\hat r^2$ of 5000 bootstraps**
```{r}
r_squared <- 
  bootstrap_results2 %>%
  select(r.squared, strap_number)
```

**Fit density plot of the product of beta estimates $\log(\hat\beta_1\times \hat\beta_2)$**

```{r}
log_betas_sd <- 
  log_betas %>%
  summarise(log_betas_sd = sd(as.numeric(log_betas),na.rm = TRUE)) %>%
  pull(log_betas_sd)

log_betas_mean <- 
  log_betas %>% 
  summarise(log_betas_mean = mean(as.numeric(log_betas), na.rm = TRUE)) %>%
  pull(log_betas_mean)

log_betas %>%
  ggplot(aes(x = log_betas)) + geom_density() +
  labs(title = "Distribution of log(Beta1 * Beta2)")
```

**Discussion:** The density curve of the the product of two beta estimates $\log(\hat\beta_1\times \hat\beta_2)$ is left-skewed. Its mean is approximately`r round(log_betas_mean,4)`, and its standard deviation is  approximately `r round(log_betas_sd,4)`.

**Fit density plot of R-Squared $\hat r^2$ **

```{r}
r_squared_sd <-
  r_squared %>%
  summarise(r_squared_sd = sd(r.squared)) %>%
  pull(r_squared_sd)

r_squared_mean <-
  r_squared %>%
  summarise(r_squared_mean = mean(r.squared)) %>%
  pull(r_squared_mean)

r_squared %>%
  ggplot(aes(x = r.squared)) + geom_density()+
  labs(title = "Distribution of R-squared")
```

**Discussion:** The density curve of R-Squared $\hat r^2$ is slightly left-skewed. Its mean is  approximately `r round(r_squared_mean,4)` and its standard deviation is  approximately `r round(r_squared_sd,4)`.

**95% CI of log of the product of beta estimates $\log(\hat\beta_1\times \hat\beta_2)$**
```{r}
CI_result <-
  log_betas %>%
  summarize(ci_lower = quantile(log_betas, 0.025, na.rm = TRUE),
            ci_upper = quantile(log_betas, 0.975, na.rm = TRUE))

CI_result_lower <- CI_result %>% pull(ci_lower)
CI_result_upper <- CI_result %>% pull(ci_upper)
```

**Discussion:** The 95% Confidence Interval of the product of beta estimates $\log(\hat\beta_1\times \hat\beta_2)$ is (`r round(CI_result_lower,4)`,`r round(CI_result_upper,4)`).

**95% CI of the R-Squared $\hat r^2$**
```{r}
CI_result2 <-
  r_squared %>%
  summarize(ci_lower = quantile(r.squared, 0.025),
            ci_upper = quantile(r.squared, 0.975)) 

CI_result_lower2 <- CI_result2 %>% pull(ci_lower)
CI_result_upper2 <- CI_result2 %>% pull(ci_upper)
```

**Discussion:** The 95% Confidence Interval of the product of beta estimates $\hat r^2$ is (`r round(CI_result_lower2,4)`,`r round(CI_result_upper2,4)`).



### Problem 3

**Import and clean dataset**
```{r}
birthweight_df = 
  read_csv("./birthweight.csv")|>
  mutate(babysex = factor(babysex), frace = factor(frace), malform = factor(malform),
         mrace = factor(mrace)) |>
  mutate(babysex = ifelse(babysex == "1", "male","female"),
         malform = ifelse(malform == "0", "absent","present"),
         frace = recode(frace, "1" = "White", "2" = "Black", "3" = "Asian", 
                        "4" = "Puerto Rican", "8" = "Other", "9" = "Unknown"),
         mrace = recode(mrace, "1" = "White", "2" = "Black", 
                        "3" = "Asian", "4" = "Puerto Rican", "8" = "Other")
         )
```
**Check for missing data**
```{r}
skimr::skim(birthweight_df)
```
**Discussion:** The cleaned dataset has no missing value. The dataset has `r ncol(birthweight_df)` variables and `r nrow(birthweight_df)` observations. The variables included in this datasets are `r names(birthweight_df)`.

**Build a regression model for birthweight**
Try to create a model with all the predictive variables available:
```{r}
model_all <- lm(bwt ~ ., data = birthweight_df)
summary(model_all)
```
**Comment:** The linear regression model with dependent variable of baby’s birth weight (grams) and independent variable of all other variablse as predictors showed some variables are statistically signification in their association with baby's birth weight. So, I will include them in my chosen model. 

The model that I choose:
```{r}
model_chosen <- lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mrace + parity + smoken, data = birthweight_df)
summary(model_chosen)
```


**Discussion:** I built a linear regression model with the independent variables (predictors) of baby's sex, baby’s head circumference at birth (centimeters), baby’s length at birth (centimeteres), mother’s weight at delivery (pounds), gestational age in weeks, mother’s race, number of live births prior to this pregnancy, and average number of cigarettes smoked per day during pregnancy, and the dependent variable of baby’s birth weight (grams). In this model, all the predictors are significantly associated with the outcome baby's birth weight except for the mother's race of asian. 


**Comparison with other models**

```{r}
model_1 <- lm(bwt ~ blength + gaweeks, data = birthweight_df)
model_2 <- lm(bwt ~ bhead + blength + babysex + bhead * blength + bhead * babysex + blength * babysex + bhead * blength * babysex, data = birthweight_df)
```

**Cross validation using `modelr`**
```{r}
cv_df = 
  crossv_mc(birthweight_df, 100) 
```


```{r}
cv_df <- 
  cv_df |> 
  mutate(
    model_chosen  = map(train, ~lm(bwt ~ babysex + bhead + blength + delwt + gaweeks + mrace + parity + smoken, data = .)),
    model_1 = map(train, ~lm(bwt ~ blength + gaweeks, data = .)),
    model_2 = map(train, ~lm(bwt ~ bhead * blength * babysex, data = .))) |>
  mutate(
    rmse_model_chosen = map2_dbl(model_chosen, test, ~rmse(model = .x, data = .y)),
    rmse_model_1 = map2_dbl(model_1, test, ~rmse(model = .x, data = .y)),
    rmse_model_2 = map2_dbl(model_2, test, ~rmse(model = .x, data = .y)))
```

**Visualize the predictive accuarcy**

```{r}
cv_df |> 
  select(starts_with("rmse")) |> 
  pivot_longer(
    everything(),
    names_to = "model", 
    values_to = "rmse",
    names_prefix = "rmse_") |> 
  mutate(model = fct_inorder(model)) |> 
  ggplot(aes(x = model, y = rmse)) + geom_violin()
```

**Discussion:** Based on the results, we can find my model has greater predictive accuracy as the rmse are lowest in my model. The model using length at birth and gestational age as predictors (main effects only) has the highest rmse, which indicates it performs worst among the three models. The model I chose has lower rmse than the model using head circumference, length, sex, and all interactions (including the three-way interaction) between these, which means my model has higher predictive accuracy than that model. 
